---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%"
)
```

# ggdatasaver

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ggdatasaver)](https://CRAN.R-project.org/package=ggdatasaver)
<!-- badges: end -->

The goal of ggdatasaver is to automatically save the data associated with your plots for you to share as supplementary material. 
Other people can then use that data instead of digitising your plots. 
Because only the data already being published as a plot is saved, there should be fewer privacy or legal complications.

## Installation

You can install the development version of ggdatasaver like so:

``` r
remotes::install_github("eliocamp/ggdatasaver")
```

## Example

ggdatasaver works automatically with knitr. 
The only thing you need to do is to define the directory where the data is saved with

```{r}
ggdatasaver::save_plot_data_in("plot-data")
```


Then, just create your ggplot2 figures as always. 
Using a chunk label is encouraged because it will be used to name the file. 

```{r mpg}
library(ggplot2)

ggplot(mtcars, aes(mpg, disp)) +
  geom_point() +
  geom_smooth()

```

After you knit, you will have a (possibly new) directory with zip files with the data of each plot. 

```{r}
fs::dir_tree("plot-data")
```

Inside that zip file there will be a csv file for each layer. 

```{r}
# Unzip the contents of mpg.zip into a temporary directory. 
dir <- file.path(tempdir(), "mpg")
utils::unzip("plot-data/mpg.zip", exdir = dir)
fs::dir_tree(dir)
```

The data of each layer is only the one used to draw the geometry. 
For example, GeomSmooth.csv has the coordinates of the fit and some other aesthetic information

```{r}
smooth <- read.csv(file.path(dir, "GeomSmooth.csv"))
knitr::kable(head(smooth))
```

And the line can be reconstructed exactly from these data. 

```{r, plot_data_dir = NULL}
ggplot(smooth, aes(x, y)) + 
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = I(fill), alpha = I(alpha))) +
  geom_line(aes(colour = I(colour), size = I(size)))
```

(Setting `plot_data_dir` to `NULL` will suppress data-saving for that chunk.)

As you can see, only the coordinates of each geom are saved, not the underlying data. 
For a more dramatic example, take this controur plot of the Old Faithful Geyser Data.


```{r faithful-density}
ggplot(faithful, aes(x = eruptions, y = waiting)) +
  geom_density_2d()
```

(Now there are two zip files in the `plot-data` directory
```{r}
fs::dir_tree("plot-data")
```
.)

ggdatasaver will save the coordinates that defined the contours, not the observations from which they were computed. 

```{r, plot_data_dir = NULL}
dir <- file.path(tempdir(), "faithful-density")
utils::unzip("plot-data/faithful-density.zip", exdir = dir)

density <- read.csv(file.path(dir, "GeomDensity2d.csv"))

ggplot(density, aes(x, y)) +
  geom_path(aes(group = group))
```

This makes it safe to share these data, as it doesn't include any more information than what's in the plot you are already sharing. 

```{r, include = FALSE}
unlink("plot-data", recursive = TRUE)
```


## Limitations


ggdatasaver has only been tested on simple plots although there's no reason it should work work with more complicated ones.
Hoever, patchwork / cowplot plots probably wont work as expected. 

For now, the package only works with one plot per chunk. 

When using ggdatasaver plots are built twice; once when saving the data and once when drawing the plot. 
This shouldn't be an issue most of the time unless your plot requires heavy computation. 
